---
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "man/figures/README-", out.width = "100%")
```

# GHSGEM: Generalized expectation-maximization (GEM) algorithm with graphical horseshoe (GHS) prior for network estimation

This R package implements a generalized expectation-maximization algorithm with a graphical horseshoe prior for network, covariance matrix, and precision matrix (inverse of the covariance) estimation.

The method is described in the article *Gene co-expression network estimation using generalized expectation-maximization algorithm with graphical horseshoe prior*. All results can be reproduced using the codes from the separate GitHub repository at <https://github.com/THautamaki/GHSGEM_results>.

## Installation

The GHSGEM package can be installed using the following code:

```{r, eval = FALSE}
if(!require("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("THautamaki/GHSGEM")
```

## Example of usage

This is a minimum working example.

```{r}
# Install R package huge if not yet installed.
if(!require("huge", quietly = TRUE)) {
  install.packages("huge")
}

library(GHSGEM)
library(huge)

n <- 200  # number of observations (sample size)
p <- 100  # number of variables
```

Generate simulated data using huge.generator.

```{r, message = FALSE, results = 'hide'}
set.seed(20250303)
sim <- huge.generator(n = n, d = p, graph = "scale-free")
```

Run the GHS GEM algorithm.

```{r}
map <- GHS_MAP_estimation(sim$data, verbose = 0)
```

Calculate and print the confusion matrix.

```{r}
(cm <- conf_matrix(sim$theta, map$Theta_est))
```

Calculate and print some performance scores.

```{r}
round(calculate_scores(cm)[, c("MCC", "F1", "TPR", "FDR")], 4)
```

Plot the true simulated network and estimated network using the R package `igraph`.

```{r, message = FALSE}
if(!require("igraph", quietly = TRUE)) {
  install.packages("igraph")
}
library(igraph)
```

Generate graph objects using `graph_from_adjacency_matrix` function from `igraph` package.

```{r}
true_network <- graph_from_adjacency_matrix(sim$theta, mode = "undirected", diag = F)
est_network <- graph_from_adjacency_matrix(map$Theta_est, mode = "undirected", diag = F)
```

Create coordinates for the nodes of the network. Use the same node placement for both plots.

```{r}
set.seed(10)
coords <- layout_with_fr(true_network)
```

Finally, plot both the true and estimated networks.

```{r network_plots, fig.dpi = 300}
par(mfrow = c(1,2), mar = c(0.1, 0.1, 1, 0.1))
plot(true_network, layout = coords, edge.width = 2, vertex.size = 4, vertex.label = NA,
     main = "True network")
plot(est_network, layout = coords, edge.width = 2, vertex.size = 4, vertex.label = NA,
     main = "Estimated network")
```
